Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control

### 2022-03-14 doc/manuscript-rmd.Rmd
slides first draft, PK&WC feedback

__Trial Design__: The design and methods of SPRINT have been published previously [@ambrosius_design_2014; @sprintFinal]. Briefly, it was a multicenter randomized clinical trial that compared two strategies for managing SBP in older adults with hypertension who were at increased risk of CVD. Participants were aged 50 years or older and had an SBP between 130 and 180 mm Hg at the screening visit, depending on the number of anti-hypertensive agents prescribed. Participants were considered to have an increased cardiovascular risk if they had clinical or subclinical cardiovascular disease, chronic kidney disease (CKD), or a Framingham Risk Score of 15% or greater or if they were aged 75 years or older. Individuals residing in a nursing home, persons with a diagnosis of dementia (based on medical record review), and those treated with medications primarily used for dementia therapy were excluded, as were persons with prevalent diabetes mellitus, history of stroke, proteinuria > 1 gram per day, or polycystic kidney disease. Individuals at 102 sites in the United States and Puerto Rico were randomized (1:1) to a SBP goal of less than 120 mm Hg (intensive treatment group, n = 4678) or a goal of less than 140 mm Hg (standard treatment group, n = 4683), using random permuted blocks with the randomization stratified by clinic site. Randomization began on November 8, 2010 and ended in March 2013. On August 20, 2015, the SPRINT trial prematurely concluded due to the clear benefit of intensive treatment. As trial follow-up visits and provision of antihypertensive medication continued through July 2016 [@sprintFinal], the current study considers a trial phase spanning from November 8, 2010 through July 2016 and an observational phase spanning from August 2017 through December, 2020 (__eFigure 1__). The study was approved by the institutional review board at each participating site, and each participant provided written informed consent. The study is registered at ClinicalTrials.gov (NCT01206062).

@@

Meta analyses of randomized trials have shown that pharmacological blood pressure (BP) lowering reduces the risk of major CVD events across the spectrum of initial BP [@rahimi_age_stratified_2021; @bundy_systolic_2017].

>>>>

__Trial Design__: The design and methods of SPRINT have been published previously [@ambrosius_design_2014; @sprintFinal]. Briefly, it was a multicenter randomized clinical trial that compared two strategies for managing SBP in older adults with hypertension who were at increased risk of CVD. Participants were aged 50 years or older and had an SBP between 130 and 180 mm Hg at the screening visit, depending on the number of anti-hypertensive agents prescribed. Participants were considered to have an increased cardiovascular risk if they had clinical or subclinical cardiovascular disease, chronic kidney disease (CKD), or a Framingham Risk Score of 15% or greater or if they were aged 75 years or older. Individuals residing in a nursing home, persons with a diagnosis of dementia (based on medical record review), and those treated with medications primarily used for dementia therapy were excluded, as were persons with prevalent diabetes mellitus, history of stroke, proteinuria > 1 gram per day, or polycystic kidney disease. Individuals at 102 sites in the United States and Puerto Rico were randomized (1:1) to a SBP goal of less than 120 mm Hg (intensive treatment group, n = 4678) or a goal of less than 140 mm Hg (standard treatment group, n = 4683), using random permuted blocks with the randomization stratified by clinic site. Randomization began on November 8, 2010 and ended in March 2013. In addition to the trial’s intervention phase, which spanned November 8, 2010 through August 20, 2015, when the administrative decision to stop the trial was made, the current study includes as part of the “trial phase” the additional period of study-provided antihypertensive medications prior to the final closeout visit on July 1, 2016 (eFigure 1). Observational follow-up continued through December 2020

As trial follow-up visits and provision of antihypertensive medication continued through July 2016 [@sprintFinal], the current study considers a trial phase spanning from November 8, 2010 through July 2016 and an observational phase spanning from August 2017 through December, 2020 (__eFigure 1__). The study was approved by the institutional review board at each participating site, and each participant provided written informed consent. The study is registered at ClinicalTrials.gov (NCT01206062).

@@

<!-- Meta analyses of randomized trials have shown that pharmacological blood pressure (BP) lowering reduces the risk of major CVD events across the spectrum of initial BP [@rahimi_age_stratified_2021; @bundy_systolic_2017]. -->

### 2022-03-14 index.Rmd
rdy to discuss slides with NMP
slides first draft, PK&WC feedback

---
title: "Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control"
subtitle: "A Secondary Analysis of SPRINT"
author: "Nicholas Pajewski, PhD"
institute: "Wake Forest School of Medicine"
date: "2016/12/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(https://upload.wikimedia.org/wikipedia/commons/b/be/Sharingan_triple.svg)

@@

tar_load(names = c(base_viz_inc,
                   base_viz_eff,
                   long_viz,
                   bp_viz,
                   ndi_baseline))

@@

@@

Notes for you go here

@@

fig$inc_cvd[[1]] + 
  theme(legend.key.size = unit(1.2, 'cm'), #change legend key size
        legend.key.height = unit(1.2, 'cm'), #change legend key height
        legend.key.width = unit(1.2, 'cm'), #change legend key width
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=16)) #change legend text font size

# If you wanted to make multiple figs on the same slide
# blank_plot <- ggplot(data.frame()) + aes() + theme_void()
# 
# legend <- get_legend(
#   fig$inc_acm[[1]]$plot + 
#     theme(legend.position = c(.55, .2),
#           legend.direction = 'horizontal')
# )
# 
# plot_grid(
#   # row 1 (legend)
#   legend, 
#   blank_plot,
#   # row 2 (figures)
#   fig$inc_acm[[1]]$plot + theme(legend.position = ''), 
#   fig$eff_acm[[1]],
#   # row 3 (table)
#   fig$inc_acm[[1]]$table, 
#   blank_plot,
#   ncol = 2,
#   align = 'v',
#   axis = 'lr',
#   rel_heights = c(1,10,2),
#   labels = c("","","A","B","",""),
#   label_x = 0.1,
#   label_y = 1.05
# )

@@

>>>>

---
title: "Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control"
subtitle: "A Secondary Analysis of SPRINT"
author: "Nicholas M. Pajewski, PhD & Byron C. Jaeger, PhD"
institute: "Wake Forest School of Medicine"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

tar_load(names = c(recoders,
                   base_viz_inc,
                   base_viz_eff,
                   bp_viz,
                   tbl_characteristics,
                   base_ds, 
                   long_ds,
                   long_inf,
                   trt_hetero,
                   base_inf,
                   ndi_baseline))

long_viz <-
  ndi_long_viz(long_inf, trt_hetero, long_ds, recoders, size_text = 4.5)

n_total <- nrow(ndi_baseline)

n_ehr <- sum(ndi_baseline$ehr_ancillary == 'Yes')

@@

## Importance and Background

- SPRINT showed that intensive treatment reduced the risk of cardiovascular and all-cause mortality.

- The trial began in 2010 and the last study closeout visit occurred in 2016.

## Objective and Methods

- We assessed whether the benefit of intensive treatment persisted after the final study closeout visit in 2016.

- We tracked cardiovascular and all-cause mortality events from the US National Death Index from 2016 to 2020.

- In a subset of participants, we examined outpatient systolic blood pressure levels measured in routine clinical practice after the trial.

@@

intensive treatment: Systolic blood pressure goal of < 120 mm Hg

legacy effect: the persistence of benefit after the trial

@@

fig$inc_cvd[[1]] +
  theme(legend.key.size = unit(1.2, 'cm'),    
        legend.key.height = unit(1.2, 'cm'),  
        legend.key.width = unit(1.2, 'cm'),   
        legend.title = element_text(size=16), 
        legend.text = element_text(size=16))

@@

bp_viz$levels + 
  theme(legend.position = 'top',
        legend.key.size = unit(1.2, 'cm'), #change legend key size
        legend.key.height = unit(1.2, 'cm'), #change legend key height
        legend.key.width = unit(1.2, 'cm'), #change legend key width
        legend.title = element_text(size=16), #change legend title font size
        legend.text = element_text(size=16))

long_viz$cvd$cohort

## Key findings

- Intensive treatment produced beneficial effects on mortality during the trial.

- The benefits associated with intensive treatment quickly attenuated as systolic blood pressure levels increased among participants who underwent intensive treatment after the trial.

- There was no evidence of sustained benefits after discontinuing the intervention protocol.

## Conclusion

Maintaining intensive BP targets throughout adulthood is likely an essential component of long-term CVD risk management.
